Algorithm explanation

### 2013-04-19 article/algo.tex
Slight fix of definition * Rename mapping on edges

\newcommand{\sembrack}[1]{[\![#1]\!}

@@

The Horn clause solving algorithm receives an input problem in Horn
graph. A Horn graph is a labelled directed graph $G=(V,E,\varphi)$
where V is a set of vertices and E is a set of edges. $\varphi: V
\rightarrow L$ is a labelling function for nodes, where L is a set of
elements of a predicate variable $P(x_1, ..., x_n)$ or a linear
inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.

@@

\begin{align*}
\mathcal{HC}_v &= \left( \bigwedge_{u \in pred(v)} \varphi(u) \right)
\Longrightarrow \varphi(v) \\ \llbracket G \rrbracket &= \bigwedge_{v
  \in V} \mathcal{HC}_v
\end{align*}

@@

Given an acyclic Horn graph (or called a Horn DAG), we define the
relation $v <^+ v'$ by
\[ v <^+ v' \Longleftrightarrow (v,v') \in E \]
By the assumption that $G$ is acyclic, $<^+$ is a well-founded
relation.  We assume without loss of generality that $V$ contains the
least element $v_\bot$ with respect to $<^+$; otherwise we can always
ignore $v_\bot$ with $\varphi(v_\bot) = 0 \leq 0$.

@@

\begin{align*}
C_v =
\begin{cases}
\left\lbrace \mathbf{a}_v = \sum_{(u,v) \in E} \mathbf{a}_u \right\rbrace \cup
\left\lbrace b_v >= \sum_{(u,v) \in E} b_u \right\rbrace \cup
\bigcup_{(u,v) \in E} C_u
& \mbox{if } \varphi(v) = P(\mathbf{x}) \vee v = v_\bot \\
\left\lbrace \mathbf{a}_v = \lambda_v \mathbf{a}_0 \right\rbrace \cup
\left\lbrace b_v = \lambda_v b_0 \right\rbrace
& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x} \leq b_0
\end{cases}
\end{align*}

@@

If $G$ has node(s) which has multiple successors, the second case is
possible. We now consider to give a solution to the such case.

Let $\hat u$ be the (least) vertex on $<^+$ with multiple successors
denoted by $succ(u)$. For each $v_i \in succ(u)$, we transform the graph
$G$ with the following operation:

\begin{align*}
\frac{
(u, v_i) \in E \wedge \varphi(u) = P(\mathbf{x}) \wedge
\Delta(u) = \left( \mathbf{a} \mathbf{x} \leq b, C \right)
}{
\begin{aligned}
E & := { (u, v_i'), (v_i', v_i) } \cup ( E \setminus (u, v_i) ) \\
\varphi(v_i') & := P_i(\mathbf{x}) \\
\Delta(v_i') & := \left( \mathbf{a}' \mathbf{x} \leq b', \textsc{Copy}(C) \right)
\end{aligned}
}
\end{align*}

@@

Note that predicate variables on greater nodes than $\hat u$ may
also have duplicated assignments, which should be combined in the
same manner as $P$ above.

>>>>

\newcommand{\sembrack}[1]{[\![#1]\!}
\newcommand{\edge}[2]{#1\rightarrow#2}
\newcommand{\path}[2]{#1\rightarrow^\star#2}

@@

The Horn clause solving algorithm receives an input problem in Horn
graph. A Horn graph is a labelled directed graph $G=(V,E,\varphi)$,
where V is a set of vertices and E is a set of labelled
edges. $\varphi: V \rightarrow L$ is a labelling function for
vertices, where L is a set of elements of a predicate variable $P(x_1,
..., x_n)$ or a linear inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
An edge from a vertex $u$ to $v$ is expressed as a 3-tuple
$(u,v,\theta)$ where $\theta$ is a finite map from variables to
variables.

@@

\begin{align*}
\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(v) \\
\llbracket G \rrbracket & = \bigwedge_{v \in V} \mathcal{HC}_v
\end{align*}

@@

Given an acyclic Horn graph (or called a Horn DAG), we define the
relation $u <^+ v$ as follows.
\[ u <^+ v \Longleftrightarrow \exists \theta. (u,v,\theta) \in E \]
By the assumption that $G$ is acyclic, $<^+$ is a well-founded
relation. We assume without loss of generality that $V$ contains the
least element $v_\bot$ with respect to $<^+$; otherwise we can always
ignore $v_\bot$ with $\varphi(v_\bot) = 0 \leq -1$.

@@

\begin{align*}
C_v =
\bigcup_{(u,v) \in E} C_u & \cup
\left\lbrace
 \mathbf{a}_v = \sum_{(u,v) \in E} \mathbf{a}_u,
 b_v \geq \sum_{(u,v) \in E} b_u
\right\rbrace \cup
\\
& \begin{cases}
\emptyset
& \mbox{if } \varphi(v) = P(\mathbf{x}) \\
\left\lbrace
 \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
 b_v = \lambda_v b_0
\right\rbrace
& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x} \leq b_0
\end{cases}
\end{align*}

@@

If $G$ has a vertex which has multiple successors, the second case is
possible. We now consider to give a solution to the such case.

Let $\hat u$ be a vertex with multiple successors denoted by
$succ(u)$. We now duplicate to relax the linear constraints.

@@

$\forall u \in V. \path{u}{v} \wedge \varphi(u) = P(\mathbf{x}).
$

Note that predicate variables on  than $\hat u$ may
also have duplicated assignments, which should be combined in the
same manner as $P$ above.
